Revisiting the Relationship between Short-Chain Fatty Acids, the Microbiota, and Colorectal Tumors

### 2019-03-20 submission/manuscript.tex
First pass at reducing the number of References

Colorectal cancer is the third leading cancer-related cause of death
within the United States (1, 2). Less than 10\% of cases can be
attributed to genetic risk factors (3). This leaves a significant role
for environmental, behavioral, and other factors such as smoking and
diet (4, 5). Colorectal cancer is thought to be initiated by a series of
mutations that accumulate as mutated cells proliferate leading to
adenomatous lesions, which are succeeded by carcinomas (3). Throughout
this progression, there are ample opportunities for bacterial
populations to create mutations, induce inflammation, and accelerate
tumorigenesis (6). Numerous studies in murine models have supported this
model (7--11). Additional cross sectional studies in humans have
identified microbiome-based biomarkers of disease (12--20). These
studies suggest that in some cases, it is the loss of bacterial
populations that produce short-chain fatty acids (SCFAs) that results in
increased inflammation and tumorigenesis.

SCFAs have have anti-inflammatory and anti-proliferative activities
(21--24). Furthermore, manipulation of SCFAs in mouse models of
colorectal cancer by direct supplementation or feeding of fiber caused
an overall reduction in tumor burden (25, 26). These results suggest
that supplementation with substrates that bacteria can ferment to
produce SCFAs may confer beneficial effects against colorectal cancer.
Regardless, there is a lack of evidence that increasing SCFA
concentrations can protect against colorectal cancer in humans. Based on
similar observations, many microbiome studies use the concentrations of
SCFAs and the presence of 16S rRNA gene sequences from organisms and the
genes involved in producing them as a biomarker of a healthy microbiota
(27--30). Case-control studies that have investigated SCFA
concentrations in colorectal cancer found that patients with carcinomas
had lower concentrations of SCFAs versus patients with adenomas or
individuals without colon tumors (31). Although this would argue that
increasing SCFA concentrations could be protective against
tumorigenesis, in randomized controlled trials fiber supplementation has
been inconsistently associated with protection against tumor formation
and recurrence (32--36). These findings temper enthusiasm for treatments
that aim to use SCFAs as biomarkers or protection against tumorigenesis.

\textbf{SCFA concentrations do not meaningfully vary with diagnosis or
treatment.} To quantify the associations between colorectal cancer, the
microbiome, and SCFAs, we quantified the concentration of acetate,
propionate, isobutyrate, and butyrate in feces of previously
characterized individuals with normal colons (N=172) and those with
colonic adenomas (N=198) or carcinomas (N=120) (14). The only SCFA that
had a significantly different concentration across the diagnoses was
isobutyrate (P=0.0091; Figure 1A). The median concentration of
isobutyrate was 3.30 mmol/kg in people with normal colons and it was
3.00 and 3.84 mmol/kg in people with colonic adenomas or carcinomas. The
difference in isobutyrate concentration between people with adenomas and
carcinomas was significantly different (P=0.0065); however, the
differences between people with normal colons and those with adenomas or
carcinomas was not significant (P=0.19 and P=0.11). The median
concentration of isobutyrate in people with normal colons (3.30 mmol/kg)
was between those with adenomas (3.00 mmol/kg) and carcinomas (3.84
mmol/kg). Among the subjects with adenomas and carcinomas, a subset
(N\textsubscript{adenoma}=41, N\textsubscript{carcinoma}=26) were
treated and sampled a year later (37). The only SCFA that changed
following treatment was isobutyrate, which decreased by 0.99 mmol/kg
(P=0.002; Figure 1B). For both the pre-treatment cross-sectional data
and the pre/post treatment data, we pooled the SCFA concentrations on a
per molecule of carbon basis and again failed to see any significant
differences (P\textgreater{}0.15). The low concentration of isobutyrate
relative to the other SCFAs, inconsistent concentrations, and unexpected
decrease in concentration with treatment makes it difficult to ascribe
much biological relevance to this observation.

\textbf{Combining SCFA and microbiome data does not improve the ability
to diagnose individual as having adenomas or carcinomas.} We previously
found that binning 16S rRNA gene sequence data into operational
taxonomic units based on 97\% similarity or into genera enabled us to
classify individuals as having adenomas or carcinomas using Random
Forest machine learning models (13, 14). We repeated that analysis but
added the concentration of the individual or pooled SCFAs as possible
features to train a model. Models trained using SCFAs to classify
individuals as having adenomas or carcinomas rather than normal colons
had median areas under the receiver operator characteristic curve
(AUROC) that were significantly greater than 0.5
(P\textsubscript{adenoma}\textless{}0.001 and
P\textsubscript{carcinoma}\textless{}0.001); however, the AUROC values
to detect the presence of adenomas or carcinomas were only 0.54 and
0.55, respectively (Figure 2A). When we trained the models with the
SCFAs concentrations and operational taxonomic unit (OTU) or genus-level
relative abundances the AUROC values were not significantly different
from the models trained without the SCFA concentrations
(P\textgreater{}0.21; Figure 2A). We also trained models using a smaller
dataset that generated shotgun metagenomic sequencing data from a subset
of our cohort (N\textsubscript{normal}=27, N\textsubscript{adenoma}=25,
and N\textsubscript{cancer}=26) (16). We binned genes extracted from the
assembled metagenomes into operational protein families (OPFs) or KEGG
categories. Again, the performance of the models trained with the
meteagenomic data did not improve when the SCFA concentrations were
added as possible features when training the model (P\textgreater{}0.24;
Figure 2A). These data demonstrate that knowledge of the SCFA profile
from a patient's fecal sample does not improve the ability to diagnose a
colonic lesion.

\textbf{Knowledge of microbial community structure does not predict SCFA
concentrations.} Regardless of a person's diagnosis, we next asked
whether the fecal community structure was predictive of fecal SCFA
concentrations. We trained Random Forest regression models using 16S
rRNA gene sequence data binned into OTUs and genera to predict the
concentration of the individual and pooled SCFAs. Regardless of the
binning method or SCFA, the largest amount of variation that the model
could explain was \textbf{0.XXXX} (Figure 2B). Next, we trained Random
Forest regression models using metagenomic sequence data binned into
OPFs and KEGG categories to predict the concentration of the individual
and pooled SCFAs. Similar to the analysis using 16S rRNA gene sequence
data, the metagenomic data was not predictive of SCFA concentration; the
largest amount of variation that the models could explain was
\textbf{0.XXXX} (Figure 2B). Because of the limited number of samples
that we were able to generate metagenomic sequence data from, we used
our 16S rRNA gene sequence data to impute metagenomes that were binned
into metabolic pathways, enzyme commission numbers, or KEGG categories
using picrust2. Again, SCFA concentrations could not be predicted based
on the imputed metagenomic data; the largest amount of variation that
the models could explain was \textbf{0.XXXX} (Figure 2B). The inability
to model SCFA concentrations from microbiome data indicates that the
knowledge of the abundance of organisms and genes is insufficient to
predict SCFA concentrations.

@@

\textbf{Study design and sampling.} The overall study design and the
resulting sequence data have been previously described (14, 37). In
brief, fecal samples were obtained from 172 individuals with normal
colons, 198 individuals with colonic adenomas, and 120 individuals with
carcinomas. Of the individuals diagnosed as having adenomas or
carcinomas, a subset (N\textsubscript{adenoma}=41 and
N\textsubscript{carcinoma}=26) were sampled after treatment of the
lesion (median=255 days between sampling, IQR=233 to 334 days). Tumor
diagnosis was made by colonoscopic examination and histopathological
review of the biopsies (14). The University of Michigan Institutional
Review Board approved the studies that generated the samples and
informed consent was obtained from all participants in accordance to the
guidelines set out by the Helsinki Declaration.

\textbf{Measuring specific SCFAs.} The measurement of acetate,
propionate, isobutyrate, and butyrate used a previously published
protocol that used High-Performance Liquid Chromatography (HPLC) (38).
Two changes were made to the protocol. First, instead of using fecal
samples suspended in DNA Genotek OmniGut tubes, we suspended frozen
fecal samples in 1 mL of PBS. Second, instead of using the average
weight of fecal sample aliquots to normalize SCFA concentrations, we
used the actual weight of the fecal samples. These methodological
changes did not affect the range of concentrations of these SCFAs
between the two studies (see Table 1 (38) and Figure 1 here).

\textbf{16S rRNA gene sequence data analysis.} Sequence data from Baxter
et al.~(14) and Sze et al.~(37) were obtained from the Sequence Read
Archive (studies SRP062005 and SRP096978) and reprocessed using using
mothur v.1.42 (39). The sequence generation workflow and processing
steps have been previously described (39, 40). In brief, the original
studies generated sequence data from V4 region of the 16S rRNA gene
using paired 250 nt reads on an Illumina MiSeq sequencer (40). The
resulting sequence data were assembled into contigs and screened for
quality aberrations including chimeras, which were removed based on the
use of UCHIME (41). Sequences were clustered into OTUs at 97\%
similarity using OptiClust (42) and assigned to the closest possible
genus using a naive Bayesian classifier with an 80\% confidence
threshold trained on the reference collection from the Ribosomal
Database Project (v.16) (43). We used PICRUSt (v.2.1.0-b) with the
recommended standard operating protocol to generate imputed metagenomes
based on the expected metabolic pathways, KEGG categories, and enzyme
commission numbers (44).

\textbf{Metagenomic DNA sequence analysis.} A subset of the samples from
the samples described by Baxter et al.~(14) were used to generate
metagenomic sequence data (N\textsubscript{normal}=27,
N\textsubscript{adenoma}=25, and N\textsubscript{cancer}=26). These data
were generated by Hannigan et al.~(16) and deposited into the Sequence
Read Archive (study SRP108915). Fecal DNA was subjected to shotgun
sequencing on an Illumina HiSeq using 125 bp paired end reads. The
archived sequences were already quality filtered and aligned to the
human genome to remove contaminating sequence data. We downloaded the
sequences and assembled them into contigs using MEGAHIT (45), which were
used to identify open reading frames (ORFs) using Prodigal (46). We
determined the abundance of each ORF by mapping the raw reads back to
the ORFs using Diamond (47). We clustered the ORFs into operational
protein families (OPFs) in which the clustered ORFs were more than 40\%
identical to each other using mmseq2 (48, 49). We also used mmseq2 to
map the ORFs to the KEGG database and clustered the ORFs according to
which category the ORFs mapped (50).

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Haggar2009}{}%
1. \textbf{Haggar F}, \textbf{Boushey R}. 2009. Colorectal cancer
epidemiology: Incidence, mortality, survival, and risk factors. Clinics
in Colon and Rectal Surgery \textbf{22}:191--197.
doi:\href{https://doi.org/10.1055/s-0029-1242458}{10.1055/s-0029-1242458}.

\leavevmode\hypertarget{ref-Siegel2016}{}%
2. \textbf{Siegel RL}, \textbf{Miller KD}, \textbf{Jemal A}. 2016.
Cancer statistics, 2016. CA: A Cancer Journal for Clinicians
\textbf{66}:7--30.
doi:\href{https://doi.org/10.3322/caac.21332}{10.3322/caac.21332}.

\leavevmode\hypertarget{ref-Fearon1990}{}%
3. \textbf{Fearon ER}, \textbf{Vogelstein B}. 1990. A genetic model for
colorectal tumorigenesis. Cell \textbf{61}:759--767.
doi:\href{https://doi.org/10.1016/0092-8674(90)90186-i}{10.1016/0092-8674(90)90186-i}.

\leavevmode\hypertarget{ref-FlissIsakov2017}{}%
4. \textbf{Fliss-Isakov N}, \textbf{Zelber-Sagi S}, \textbf{Webb M},
\textbf{Halpern Z}, \textbf{Kariv R}. 2017. Smoking habits are strongly
associated with colorectal polyps in a population-based case-control
study. Journal of Clinical Gastroenterology 1.
doi:\href{https://doi.org/10.1097/mcg.0000000000000935}{10.1097/mcg.0000000000000935}.

\leavevmode\hypertarget{ref-Lee2015}{}%
5. \textbf{Lee J}, \textbf{Jeon JY}, \textbf{Meyerhardt JA}. 2015. Diet
and lifestyle in survivors of colorectal cancer. Hematology/Oncology
Clinics of North America \textbf{29}:1--27.
doi:\href{https://doi.org/10.1016/j.hoc.2014.09.005}{10.1016/j.hoc.2014.09.005}.

\leavevmode\hypertarget{ref-Flynn2016}{}%
6. \textbf{Flynn KJ}, \textbf{Baxter NT}, \textbf{Schloss PD}. 2016.
Metabolic and community synergy of oral bacteria in colorectal cancer.
mSphere \textbf{1}.
doi:\href{https://doi.org/10.1128/msphere.00102-16}{10.1128/msphere.00102-16}.

\leavevmode\hypertarget{ref-Zackular2013}{}%
7. \textbf{Zackular JP}, \textbf{Baxter NT}, \textbf{Iverson KD},
\textbf{Sadler WD}, \textbf{Petrosino JF}, \textbf{Chen GY},
\textbf{Schloss PD}. 2013. The gut microbiome modulates colon
tumorigenesis. mBio \textbf{4}:e00692--13--e00692--13.
doi:\href{https://doi.org/10.1128/mbio.00692-13}{10.1128/mbio.00692-13}.

\leavevmode\hypertarget{ref-Baxter2014}{}%
8. \textbf{Baxter NT}, \textbf{Zackular JP}, \textbf{Chen GY},
\textbf{Schloss PD}. 2014. Structure of the gut microbiome following
colonization with human feces determines colonic tumor burden.
Microbiome \textbf{2}:20.
doi:\href{https://doi.org/10.1186/2049-2618-2-20}{10.1186/2049-2618-2-20}.

\leavevmode\hypertarget{ref-Zackular2015}{}%
9. \textbf{Zackular JP}, \textbf{Baxter NT}, \textbf{Chen GY},
\textbf{Schloss PD}. 2015. Manipulation of the gut microbiota reveals
role in colon tumorigenesis. mSphere \textbf{1}:e00001--15.
doi:\href{https://doi.org/10.1128/msphere.00001-15}{10.1128/msphere.00001-15}.

\leavevmode\hypertarget{ref-DeStefanoShields2016}{}%
10. \textbf{Shields CED}, \textbf{Meerbeke SWV}, \textbf{Housseau F},
\textbf{Wang H}, \textbf{Huso DL}, \textbf{Casero RA}, \textbf{O'Hagan
HM}, \textbf{Sears CL}. 2016. Reduction of murine colon tumorigenesis
driven by EnterotoxigenicBacteroides fragilisUsing cefoxitin treatment.
Journal of Infectious Diseases \textbf{214}:122--129.
doi:\href{https://doi.org/10.1093/infdis/jiw069}{10.1093/infdis/jiw069}.

\leavevmode\hypertarget{ref-Tomkovich2017}{}%
11. \textbf{Tomkovich S}, \textbf{Yang Y}, \textbf{Winglee K},
\textbf{Gauthier J}, \textbf{Mühlbauer M}, \textbf{Sun X},
\textbf{Mohamadzadeh M}, \textbf{Liu X}, \textbf{Martin P}, \textbf{Wang
GP}, \textbf{Oswald E}, \textbf{Fodor AA}, \textbf{Jobin C}. 2017.
Locoregional effects of microbiota in a preclinical model of colon
carcinogenesis. Cancer Research \textbf{77}:2620--2632.
doi:\href{https://doi.org/10.1158/0008-5472.can-16-3472}{10.1158/0008-5472.can-16-3472}.

\leavevmode\hypertarget{ref-Kostic2011}{}%
12. \textbf{Kostic AD}, \textbf{Gevers D}, \textbf{Pedamallu CS},
\textbf{Michaud M}, \textbf{Duke F}, \textbf{Earl AM}, \textbf{Ojesina
AI}, \textbf{Jung J}, \textbf{Bass AJ}, \textbf{Tabernero J},
\textbf{Baselga J}, \textbf{Liu C}, \textbf{Shivdasani RA},
\textbf{Ogino S}, \textbf{Birren BW}, \textbf{Huttenhower C},
\textbf{Garrett WS}, \textbf{Meyerson M}. 2011. Genomic analysis
identifies association of fusobacterium with colorectal carcinoma.
Genome Research \textbf{22}:292--298.
doi:\href{https://doi.org/10.1101/gr.126573.111}{10.1101/gr.126573.111}.

\leavevmode\hypertarget{ref-Sze2018}{}%
13. \textbf{Sze MA}, \textbf{Schloss PD}. 2018. Leveraging existing 16S
rRNA gene surveys to identify reproducible biomarkers in individuals
with colorectal tumors.
doi:\href{https://doi.org/10.1101/285486}{10.1101/285486}.

\leavevmode\hypertarget{ref-Baxter2016}{}%
14. \textbf{Baxter NT}, \textbf{Ruffin MT}, \textbf{Rogers MAM},
\textbf{Schloss PD}. 2016. Microbiota-based model improves the
sensitivity of fecal immunochemical test for detecting colonic lesions.
Genome Medicine \textbf{8}.
doi:\href{https://doi.org/10.1186/s13073-016-0290-3}{10.1186/s13073-016-0290-3}.

\leavevmode\hypertarget{ref-Zackular2014}{}%
15. \textbf{Zackular JP}, \textbf{Rogers MAM}, \textbf{Ruffin MT},
\textbf{Schloss PD}. 2014. The human gut microbiome as a screening tool
for colorectal cancer. Cancer Prevention Research \textbf{7}:1112--1121.
doi:\href{https://doi.org/10.1158/1940-6207.capr-14-0129}{10.1158/1940-6207.capr-14-0129}.

\leavevmode\hypertarget{ref-Hannigan2017}{}%
16. \textbf{Hannigan GD}, \textbf{Duhaime MB}, \textbf{Ruffin MT},
\textbf{Koumpouras CC}, \textbf{Schloss PD}. 2017. Diagnostic potential
\& the interactive dynamics of the colorectal cancer virome.
doi:\href{https://doi.org/10.1101/152868}{10.1101/152868}.

>>>>

Colorectal cancer is the third leading cancer-related cause of death
within the United States (1). Less than 10\% of cases can be attributed
to genetic risk factors (2). This leaves a significant role for
environmental, behavioral, and other factors such as smoking and diet
(3, 4). Colorectal cancer is thought to be initiated by a series of
mutations that accumulate as mutated cells proliferate leading to
adenomatous lesions, which are succeeded by carcinomas (2). Throughout
this progression, there are ample opportunities for bacterial
populations to create mutations, induce inflammation, and accelerate
tumorigenesis. Numerous studies in murine models have supported this
model (5--7). Additional cross sectional studies in humans have
identified microbiome-based biomarkers of disease (8). These studies
suggest that in some cases, it is the loss of bacterial populations that
produce short-chain fatty acids (SCFAs) that results in increased
inflammation and tumorigenesis.

SCFAs have have anti-inflammatory and anti-proliferative activities (9).
Manipulation of SCFAs in mouse models of colorectal cancer by direct
supplementation or feeding of fiber caused an overall reduction in tumor
burden (10). These results suggest that supplementation with substrates
that bacteria can ferment to produce SCFAs may confer beneficial effects
against colorectal cancer. Regardless, there is a lack of evidence that
increasing SCFA concentrations can protect against colorectal cancer in
humans. Based on similar observations, many microbiome studies use the
concentrations of SCFAs and the presence of 16S rRNA gene sequences from
organisms and the genes involved in producing them as a biomarker of a
healthy microbiota (11, 12). Case-control studies that have investigated
SCFA concentrations in colorectal cancer found that patients with
carcinomas had lower concentrations of SCFAs versus patients with
adenomas or individuals without colon tumors (13). Although this would
argue that increasing SCFA concentrations could be protective against
tumorigenesis, in randomized controlled trials fiber supplementation has
been inconsistently associated with protection against tumor formation
and recurrence (14, 15). These findings temper enthusiasm for treatments
that aim to use SCFAs as biomarkers or protection against tumorigenesis.

\textbf{SCFA concentrations do not meaningfully vary with diagnosis or
treatment.} To quantify the associations between colorectal cancer, the
microbiome, and SCFAs, we quantified the concentration of acetate,
propionate, isobutyrate, and butyrate in feces of previously
characterized individuals with normal colons (N=172) and those with
colonic adenomas (N=198) or carcinomas (N=120) (16). The only SCFA that
had a significantly different concentration across the diagnoses was
isobutyrate (P=0.0091; Figure 1A). Interestingly, the median
concentration of isobutyrate was 3.30 mmol/kg in people with normal
colons and it was 3.00 and 3.84 mmol/kg in people with colonic adenomas
or carcinomas, respectively. The difference in isobutyrate concentration
between people with adenomas and carcinomas was significantly different
(P=0.0065); however, the differences between people with normal colons
and those with adenomas or carcinomas was not significant (P=0.19 and
P=0.11). Among the subjects with adenomas and carcinomas, a subset
(N\textsubscript{adenoma}=41, N\textsubscript{carcinoma}=26) were
treated and sampled a year later (17). The only SCFA that changed
following treatment was isobutyrate, which decreased by 0.99 mmol/kg
(P=0.002; Figure 1B). For both the pre-treatment cross-sectional data
and the pre/post treatment data, we pooled the SCFA concentrations on a
per molecule of carbon basis and again failed to see any significant
differences (P\textgreater{}0.15). The low concentration of isobutyrate
relative to the other SCFAs, inconsistent concentrations, and unexpected
decrease in concentration with treatment makes it difficult to ascribe
much biological relevance to this observation.

\textbf{Combining SCFA and microbiome data does not improve the ability
to diagnose individual as having adenomas or carcinomas.} We previously
found that binning 16S rRNA gene sequence data into operational
taxonomic units based on 97\% similarity or into genera enabled us to
classify individuals as having adenomas or carcinomas using Random
Forest machine learning models (8, 16). We repeated that analysis but
added the concentration of the individual or pooled SCFAs as possible
features to train a model. Models trained using SCFAs to classify
individuals as having adenomas or carcinomas rather than normal colons
had median areas under the receiver operator characteristic curve
(AUROC) that were significantly greater than 0.5
(P\textsubscript{adenoma}\textless{}0.001 and
P\textsubscript{carcinoma}\textless{}0.001); however, the AUROC values
to detect the presence of adenomas or carcinomas were only 0.54 and
0.55, respectively (Figure 2A). When we trained the models with the
SCFAs concentrations and operational taxonomic unit (OTU) or genus-level
relative abundances the AUROC values were not significantly different
from the models trained without the SCFA concentrations
(P\textgreater{}0.21; Figure 2A). We also trained models using a smaller
dataset that generated shotgun metagenomic sequencing data from a subset
of our cohort (N\textsubscript{normal}=27, N\textsubscript{adenoma}=25,
and N\textsubscript{cancer}=26) (18). We binned genes extracted from the
assembled metagenomes into operational protein families (OPFs) or KEGG
categories. Again, the performance of the models trained with the
meteagenomic data did not improve when the SCFA concentrations were
added as possible features when training the model (P\textgreater{}0.24;
Figure 2A). These data demonstrate that knowledge of the SCFA profile
from a patient's fecal sample does not improve the ability to diagnose a
colonic lesion.

\textbf{Knowledge of microbial community structure does not predict SCFA
concentrations.} Regardless of a person's diagnosis, we next asked
whether the fecal community structure was predictive of fecal SCFA
concentrations. We trained Random Forest regression models using 16S
rRNA gene sequence data binned into OTUs and genera to predict the
concentration of the individual and pooled SCFAs. Regardless of the
binning method or SCFA, the largest amount of variation that the model
could explain was \textbf{0.XXXX} (Figure 2B). Next, we trained Random
Forest regression models using metagenomic sequence data binned into
OPFs and KEGG categories to predict the concentration of the individual
and pooled SCFAs. Similar to the analysis using 16S rRNA gene sequence
data, the metagenomic data was not predictive of SCFA concentration; the
largest amount of variation that the models could explain was
\textbf{0.XXXX} (Figure 2B). Because of the limited number of samples
that we were able to generate metagenomic sequence data from, we used
our 16S rRNA gene sequence data to impute metagenomes that were binned
into metabolic pathways, enzyme commission numbers, or KEGG categories
using PICRUSt. Again, SCFA concentrations could not be predicted based
on the imputed metagenomic data; the largest amount of variation that
the models could explain was \textbf{0.XXXX} (Figure 2B). The inability
to model SCFA concentrations from microbiome data indicates that the
knowledge of the abundance of organisms and genes is insufficient to
predict SCFA concentrations.

@@

\textbf{Study design and sampling.} The overall study design and the
resulting sequence data have been previously described (16, 17). In
brief, fecal samples were obtained from 172 individuals with normal
colons, 198 individuals with colonic adenomas, and 120 individuals with
carcinomas. Of the individuals diagnosed as having adenomas or
carcinomas, a subset (N\textsubscript{adenoma}=41 and
N\textsubscript{carcinoma}=26) were sampled after treatment of the
lesion (median=255 days between sampling, IQR=233 to 334 days). Tumor
diagnosis was made by colonoscopic examination and histopathological
review of the biopsies (16). The University of Michigan Institutional
Review Board approved the studies that generated the samples and
informed consent was obtained from all participants in accordance to the
guidelines set out by the Helsinki Declaration.

\textbf{Measuring specific SCFAs.} The measurement of acetate,
propionate, isobutyrate, and butyrate used a previously published
protocol that used High-Performance Liquid Chromatography (HPLC) (19).
Two changes were made to the protocol. First, instead of using fecal
samples suspended in DNA Genotek OmniGut tubes, we suspended frozen
fecal samples in 1 mL of PBS. Second, instead of using the average
weight of fecal sample aliquots to normalize SCFA concentrations, we
used the actual weight of the fecal samples. These methodological
changes did not affect the range of concentrations of these SCFAs
between the two studies.

\textbf{16S rRNA gene sequence data analysis.} Sequence data from Baxter
et al.~(16) and Sze et al.~(17) were obtained from the Sequence Read
Archive (studies SRP062005 and SRP096978) and reprocessed using using
mothur v.1.42 (20). The original studies generated sequence data from V4
region of the 16S rRNA gene using paired 250 nt reads on an Illumina
MiSeq sequencer. The resulting sequence data were assembled into
contigs, screened to remove low quality contigs and chimeras. The
curated sequences were then clustered into OTUs at a 97\% similarity
threshold and assigned to the closest possible genus with an 80\%
confidence threshold trained on the reference collection from the
Ribosomal Database Project (v.16). We used PICRUSt (v.2.1.0-b) with the
recommended standard operating protocol to generate imputed metagenomes
based on the expected metabolic pathways, KEGG categories, and enzyme
commission numbers (21).

\textbf{Metagenomic DNA sequence analysis.} A subset of the samples from
the samples described by Baxter et al.~(16) were used to generate
metagenomic sequence data (N\textsubscript{normal}=27,
N\textsubscript{adenoma}=25, and N\textsubscript{cancer}=26). These data
were generated by Hannigan et al.~(18) and deposited into the Sequence
Read Archive (study SRP108915). Fecal DNA was subjected to shotgun
sequencing on an Illumina HiSeq using 125 bp paired end reads. The
archived sequences were already quality filtered and aligned to the
human genome to remove contaminating sequence data. We downloaded the
sequences and assembled them into contigs using MEGAHIT (22), which were
used to identify open reading frames (ORFs) using Prodigal (23). We
determined the abundance of each ORF by mapping the raw reads back to
the ORFs using Diamond (24). We clustered the ORFs into operational
protein families (OPFs) in which the clustered ORFs were more than 40\%
identical to each other using mmseq2 (25). We also used mmseq2 to map
the ORFs to the KEGG database and clustered the ORFs according to which
category the ORFs mapped.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Siegel2016}{}%
1. \textbf{Siegel RL}, \textbf{Miller KD}, \textbf{Jemal A}. 2016.
Cancer statistics, 2016. CA: A Cancer Journal for Clinicians
\textbf{66}:7--30.
doi:\href{https://doi.org/10.3322/caac.21332}{10.3322/caac.21332}.

\leavevmode\hypertarget{ref-Fearon1990}{}%
2. \textbf{Fearon ER}, \textbf{Vogelstein B}. 1990. A genetic model for
colorectal tumorigenesis. Cell \textbf{61}:759--767.
doi:\href{https://doi.org/10.1016/0092-8674(90)90186-i}{10.1016/0092-8674(90)90186-i}.

\leavevmode\hypertarget{ref-FlissIsakov2017}{}%
3. \textbf{Fliss-Isakov N}, \textbf{Zelber-Sagi S}, \textbf{Webb M},
\textbf{Halpern Z}, \textbf{Kariv R}. 2017. Smoking habits are strongly
associated with colorectal polyps in a population-based case-control
study. Journal of Clinical Gastroenterology 1.
doi:\href{https://doi.org/10.1097/mcg.0000000000000935}{10.1097/mcg.0000000000000935}.

\leavevmode\hypertarget{ref-Lee2015}{}%
4. \textbf{Lee J}, \textbf{Jeon JY}, \textbf{Meyerhardt JA}. 2015. Diet
and lifestyle in survivors of colorectal cancer. Hematology/Oncology
Clinics of North America \textbf{29}:1--27.
doi:\href{https://doi.org/10.1016/j.hoc.2014.09.005}{10.1016/j.hoc.2014.09.005}.

\leavevmode\hypertarget{ref-Zackular2013}{}%
5. \textbf{Zackular JP}, \textbf{Baxter NT}, \textbf{Iverson KD},
\textbf{Sadler WD}, \textbf{Petrosino JF}, \textbf{Chen GY},
\textbf{Schloss PD}. 2013. The gut microbiome modulates colon
tumorigenesis. mBio \textbf{4}:e00692--13--e00692--13.
doi:\href{https://doi.org/10.1128/mbio.00692-13}{10.1128/mbio.00692-13}.

\leavevmode\hypertarget{ref-DeStefanoShields2016}{}%
6. \textbf{Shields CED}, \textbf{Meerbeke SWV}, \textbf{Housseau F},
\textbf{Wang H}, \textbf{Huso DL}, \textbf{Casero RA}, \textbf{O'Hagan
HM}, \textbf{Sears CL}. 2016. Reduction of murine colon tumorigenesis
driven by Enterotoxigenic Bacteroides fragilis using cefoxitin
treatment. Journal of Infectious Diseases \textbf{214}:122--129.
doi:\href{https://doi.org/10.1093/infdis/jiw069}{10.1093/infdis/jiw069}.

\leavevmode\hypertarget{ref-Tomkovich2017}{}%
7. \textbf{Tomkovich S}, \textbf{Yang Y}, \textbf{Winglee K},
\textbf{Gauthier J}, \textbf{Mühlbauer M}, \textbf{Sun X},
\textbf{Mohamadzadeh M}, \textbf{Liu X}, \textbf{Martin P}, \textbf{Wang
GP}, \textbf{Oswald E}, \textbf{Fodor AA}, \textbf{Jobin C}. 2017.
Locoregional effects of microbiota in a preclinical model of colon
carcinogenesis. Cancer Research \textbf{77}:2620--2632.
doi:\href{https://doi.org/10.1158/0008-5472.can-16-3472}{10.1158/0008-5472.can-16-3472}.

\leavevmode\hypertarget{ref-Sze2018}{}%
8. \textbf{Sze MA}, \textbf{Schloss PD}. 2018. Leveraging existing 16S
rRNA gene surveys to identify reproducible biomarkers in individuals
with colorectal tumors.
doi:\href{https://doi.org/10.1101/285486}{10.1101/285486}.

\leavevmode\hypertarget{ref-OKeefe2016}{}%
9. \textbf{O'Keefe SJD}. 2016. Diet, microorganisms and their
metabolites and colon cancer. Nature Reviews Gastroenterology \&
Hepatology \textbf{13}:691--706.
doi:\href{https://doi.org/10.1038/nrgastro.2016.165}{10.1038/nrgastro.2016.165}.

\leavevmode\hypertarget{ref-Bishehsari2018}{}%
10. \textbf{Bishehsari F}, \textbf{Engen P}, \textbf{Preite N},
\textbf{Tuncil Y}, \textbf{Naqib A}, \textbf{Shaikh M}, \textbf{Rossi
M}, \textbf{Wilber S}, \textbf{Green S}, \textbf{Hamaker B},
\textbf{Khazaie K}, \textbf{Voigt R}, \textbf{Forsyth C},
\textbf{Keshavarzian A}. 2018. Dietary fiber treatment corrects the
composition of gut microbiota, promotes SCFA production, and suppresses
colon carcinogenesis. Genes \textbf{9}:102.
doi:\href{https://doi.org/10.3390/genes9020102}{10.3390/genes9020102}.

\leavevmode\hypertarget{ref-Sanna2019}{}%
11. \textbf{Sanna S}, \textbf{Zuydam NR van}, \textbf{Mahajan A},
\textbf{Kurilshikov A}, \textbf{Vila AV}, \textbf{Võsa U},
\textbf{Mujagic Z}, \textbf{Masclee AAM}, \textbf{Jonkers DMAE},
\textbf{Oosting M}, \textbf{Joosten LAB}, \textbf{Netea MG},
\textbf{Franke L}, \textbf{Zhernakova A}, \textbf{Fu J},
\textbf{Wijmenga C}, \textbf{McCarthy MI}. 2019. Causal relationships
among the gut microbiome, short-chain fatty acids and metabolic
diseases. Nature Genetics.
doi:\href{https://doi.org/10.1038/s41588-019-0350-x}{10.1038/s41588-019-0350-x}.

\leavevmode\hypertarget{ref-Meisel2016}{}%
12. \textbf{Meisel M}, \textbf{Mayassi T}, \textbf{Fehlner-Peach H},
\textbf{Koval JC}, \textbf{O'Brien SL}, \textbf{Hinterleitner R},
\textbf{Lesko K}, \textbf{Kim S}, \textbf{Bouziat R}, \textbf{Chen L},
\textbf{Weber CR}, \textbf{Mazmanian SK}, \textbf{Jabri B},
\textbf{Antonopoulos DA}. 2016. Interleukin-15 promotes intestinal
dysbiosis with butyrate deficiency associated with increased
susceptibility to colitis. The ISME Journal \textbf{11}:15--30.
doi:\href{https://doi.org/10.1038/ismej.2016.114}{10.1038/ismej.2016.114}.

\leavevmode\hypertarget{ref-Ohigashi2013}{}%
13. \textbf{Ohigashi S}, \textbf{Sudo K}, \textbf{Kobayashi D},
\textbf{Takahashi O}, \textbf{Takahashi T}, \textbf{Asahara T},
\textbf{Nomoto K}, \textbf{Onodera H}. 2013. Changes of the intestinal
microbiota, short chain fatty acids, and fecal pH in patients with
colorectal cancer. Digestive Diseases and Sciences
\textbf{58}:1717--1726.
doi:\href{https://doi.org/10.1007/s10620-012-2526-4}{10.1007/s10620-012-2526-4}.

\leavevmode\hypertarget{ref-Yao2017}{}%
14. \textbf{Yao Y}, \textbf{Suo T}, \textbf{Andersson R}, \textbf{Cao
Y}, \textbf{Wang C}, \textbf{Lu J}, \textbf{Chui E}. 2017. Dietary fibre
for the prevention of recurrent colorectal adenomas and carcinomas.
Cochrane Database of Systematic Reviews.
doi:\href{https://doi.org/10.1002/14651858.cd003430.pub2}{10.1002/14651858.cd003430.pub2}.

\leavevmode\hypertarget{ref-Gianfredi2018}{}%
15. \textbf{Gianfredi V}, \textbf{Salvatori T}, \textbf{Villarini M},
\textbf{Moretti M}, \textbf{Nucci D}, \textbf{Realdon S}. 2018. Is
dietary fibre truly protective against colon cancer? A systematic review
and meta-analysis. International Journal of Food Sciences and Nutrition
\textbf{69}:904--915.
doi:\href{https://doi.org/10.1080/09637486.2018.1446917}{10.1080/09637486.2018.1446917}.

\leavevmode\hypertarget{ref-Baxter2016}{}%
16. \textbf{Baxter NT}, \textbf{Ruffin MT}, \textbf{Rogers MAM},
\textbf{Schloss PD}. 2016. Microbiota-based model improves the
sensitivity of fecal immunochemical test for detecting colonic lesions.
Genome Medicine \textbf{8}.
doi:\href{https://doi.org/10.1186/s13073-016-0290-3}{10.1186/s13073-016-0290-3}.
